---
title: "HW6 : Outlier Analysis"
author: "Apoorva Srinivasa, Julia Tcholakova "
date: "3/4/2019"
output: pdf_document
---


# Code for regression and resulting model.


```{r }
housing_data <- read.csv("housing.data.csv", header = FALSE)
colnames(housing_data) <- c("CRIM","ZN","INDUS","CHAS" ,"NOX" , "RM" , "AGE" , "DIS" ,"RAD" ,
                            "TAX" ,"PTRATIO", "B" , "LSTAT" , "MEDV" )
fit <- lm(MEDV~ ., data =housing_data)
summary(fit)
```

\pagebreak

# Diagnostic plot 

```{r pressure,fig.height= 7, echo=FALSE}
par(mfrow=c(2,2))
plot(fit)
```



```
On closely observing the Std Residuals Vs Fitted plot, we see that the point indexes - 369, 373, 372 
have standardized residuals that are more than 4 std.deviations away from the mean and also high 
cook's distance. Additionally, the rows 366, 381 have both high Cook's distance (Those points that 
are greater than 4/length( all cooksdistances)) as well as high leverage(Those points that have 
more than 3 times the mean of the leverages), which is not favourable . Hence we will consider 
these 5 points as outliers and remove them from our dataset. After removing the five points 
that have questionably high standardized residuals, levergae and Cook's distance, we build 
a new model and observe the resulting plots.
On examining the plots from the new model, we find more point indexes, 368, 370, 371 and 413, have
very high Standarized residuals and Cook's distance. Hence the final model is built after removing
all these observations.
```

\pagebreak

# New Diagnostic plot 

```{r fig.height= 7, echo=FALSE}
##### Checking for outliers using leverage, Cook's distance and Standardized residuals
# Leverage cut off 3 times mean value
high_lev <- as.numeric(names(hatvalues(fit)[hatvalues(fit) > 3 * mean(hatvalues(fit))]))
# Std residuals
std.res <- rstandard(fit)[abs(rstandard(fit)) > 4]
possible_outliers <- as.numeric(names(std.res))
# Cooks distance cut off greater than 4
high_cooks <- as.numeric(names(cooks.distance(fit)[cooks.distance(fit) > 4 / length(cooks.distance(fit))]))
# Remove the points which have both high leverage and high Cook's distance
high_lev_cooks <- high_lev[high_lev %in% high_cooks]
possible_outliers <- c(possible_outliers,high_lev_cooks)

outlier_treated_housing_data <- housing_data[-c(possible_outliers),]
outlier_fit <- lm(MEDV~ ., data =outlier_treated_housing_data)


std.res1 <- rstandard(outlier_fit)[abs(rstandard(outlier_fit)) > 4]
possible_outliers1 <- as.numeric(names(std.res1))
outlier_treated_housing_data <- housing_data[-c(possible_outliers,possible_outliers1),]
outlier_fit <- lm(MEDV~ ., data =outlier_treated_housing_data)

# Diagnistic Plot
par(mfrow=c(2,2))
plot(outlier_fit)
```

\pagebreak

# Code for regression model after removing outliers

```{r fig.height= 7, eval = FALSE}
# Checking for outliers using leverage, Cook's distance and Standardized residuals
# Leverage cut off 3 times mean value
high_lev <- as.numeric(names(hatvalues(fit)[hatvalues(fit) > 3 * mean(hatvalues(fit))]))
# Std residuals
std.res <- rstandard(fit)[abs(rstandard(fit)) > 4]
possible_outliers <- as.numeric(names(std.res))
# Cooks distance cut off greater than 4
high_cooks <- as.numeric(names(cooks.distance(fit)[cooks.distance(fit) > 4 / 
                                                     length(cooks.distance(fit))]))
# Points which have both high leverage and high Cook's distance
high_lev_cooks <- high_lev[high_lev %in% high_cooks]
possible_outliers <- c(possible_outliers,high_lev_cooks)


outlier_treated_housing_data <- housing_data[-c(possible_outliers),]
outlier_fit <- lm(MEDV~ ., data =outlier_treated_housing_data)

par(mfrow=c(2,2))
plot(outlier_fit)

# Repeating outlier removal step
std.res1 <- rstandard(outlier_fit)[abs(rstandard(outlier_fit)) > 4]
possible_outliers1 <- as.numeric(names(std.res1))
outlier_treated_housing_data <- housing_data[-c(possible_outliers,possible_outliers1),]
outlier_fit <- lm(MEDV~ ., data =outlier_treated_housing_data)
possible_outliers1

# Diagnistic Plot
par(mfrow=c(2,2))
plot(outlier_fit)
```

\pagebreak

# Box_Cox transformation Plot and choosing best value for Lamda

```{r , echo=FALSE}
library(MASS)
library(faraway)
bc <- boxcox(outlier_fit, plotit = TRUE)
#Code to obtain best lamda
best_lamda <- bc$x[which(bc$y ==max(bc$y))] 
```

##  The best value of Lamda is  `r best_lamda` 

\pagebreak

# Standardized Residuals Plot and True Vs Fitted plot after Box-Cox

```{r , echo=FALSE}
boxcox_fit <- lm ((( MEDV ^ best_lamda) - 1 )/ best_lamda ~ ., data = outlier_treated_housing_data)
# Standardized Residuals Vs Fitted
reverse_transformed_y <- ((boxcox_fit$fitted.values *best_lamda)+1)^(1/best_lamda)
plot(reverse_transformed_y, rstandard(boxcox_fit), ylab="Standardized Residuals", xlab="Fited Value", main="Processed Data")
abline(0, 0, col='red')    


plot(reverse_transformed_y, outlier_treated_housing_data$MEDV, main ="Fitted Vs True House Price", xlab="Fitted House Price", ylab="True House Price") 
abline(1,1, col='red')    

```

\pagebreak

# Code for Box-Cox transformation and regression

```{r fig.height= 7, eval = FALSE}

bc <- boxcox(outlier_fit, plotit = TRUE)
#Code to obtain best lamda
best_lamda <- bc$x[which(bc$y ==max(bc$y))] 

boxcox_fit <- lm ((( MEDV ^ best_lamda) - 1 )/ best_lamda ~ ., 
                  data = outlier_treated_housing_data)

# Standardized Residuals Vs Fitted
reverse_transformed_y <- ((boxcox_fit$fitted.values *best_lamda)+1)^(1/best_lamda)
plot(reverse_transformed_y, rstandard(boxcox_fit), ylab="Standardized Residuals",
     xlab="Fited Value", main="Processed Data")
abline(0, 0, col='red')    

#True Vs Fitted Plot
plot(reverse_transformed_y, outlier_treated_housing_data$MEDV, 
     main ="Fitted Vs True House Price", xlab="Fitted House Price", ylab="True House Price") 
abline(1,1, col='red') 
```




